---
---
date: "`r format(Sys.time(), '%d %B, %Y')`" \
author: "Joselynn Wallace" \
output: pdf_document 

#https://garsia.math.yorku.ca/~zabrocki/latexpanel/latexpanel.pdf

# Problem Set-LE

## Problem 1
Explain how to solve the following systems, and write the solution set using set-builder notation. For (a), perform all row-reduction steps by hand; for (b) and (c), use sage/octave to do the row-reduction for you. 

## Problem 1A.

**Starting Equation**

$$
\begin{equation}
x_1 - x_2 + 3x_3 = 1 \\
-4x_1 + 2x_2 + x_3 = -1 \\
10x_1 - 4x_2 + 5x_3 = 1
\end{equation}
$$
Convert this system of linear equations to an augmented matrix:

**Starting Matrix:** \

$$
\begin{bmatrix} 
3  & -1 & 3 & 1 \\ 
-4 & 2 & 1 & -1 \\
10 & -4 & 5 & 1 \\
\end{bmatrix}
$$

Calculate the RREF:
  
```{r include = F}
i = 0
```
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 1 with the sum of row 1 and row 2 times $\frac{1}{2}$ **(R1 = R1 + $\frac{1}{2}$R2)** \
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & \frac{1}{2} \\ 
-4 & 2 & 1 & -1 \\
10 & -4 & 5 & 1 \\
\end{bmatrix}
$$
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 2 with the sum of row 2 and 4 times row 1. **(R2 = R2 + 4R1)**
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & \frac{1}{2} \\ 
0 & 2 & 15 & 1 \\
10 & -4 & 5 & 1 \\
\end{bmatrix}
$$
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 3 with the difference between row 3 and row 1 times 10. **(R3 = R3 - 10R1)**
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & \frac{1}{2} \\ 
0 & 2 & 15 & 1 \\
0 & -4 & -30 & -4 \\
\end{bmatrix}
$$
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 3 with the sum of row 3 and row 2 times 2. **(R3 = R3 + 2R2)**
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & \frac{1}{2} \\ 
0 & 2 & 15 & 1 \\
0 & 0 & 0 & -2 \\
\end{bmatrix}
$$
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 3 with row 3 times -1. **(R3 = -1R3)**
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & \frac{1}{2} \\ 
0 & 2 & 15 & 1 \\
0 & 0 & 0 & 2 \\
\end{bmatrix}
$$
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 3 with row 3 times $\frac{1}{2}$ **(R3 = $\frac{1}{2}$R3)**
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & \frac{1}{2} \\ 
0 & 2 & 15 & 1 \\
0 & 0 & 0 & 1 \\
\end{bmatrix}
$$
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 2 with the difference between row 2 and row 3 **(R2 = R2 - R3)**
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & \frac{1}{2} \\ 
0 & 2 & 15 & 0 \\
0 & 0 & 0 & 1 \\
\end{bmatrix}
$$

```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 1 with row 1 - $\frac{1}{2}$R3. **(R1 = R1 - $\frac{1}{2}$R3)**
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & 0 \\ 
0 & 2 & 15 & 0 \\
0 & 0 & 0 & 1 \\
\end{bmatrix}
$$
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 2 with row 2 times $\frac{1}{2}$ and **this matrix is in reduced row echelon form.**. **(R2 = $\frac{1}{2}$R2)**
$$
\begin{bmatrix} 
1  & 0 & 3\frac{1}{2} & 0 \\ 
0 & 1 & 7\frac{1}{2} & 0 \\
0 & 0 & 0 & 1 \\
\end{bmatrix}
$$

The last line is essentially saying that 0+0+0 = 1, which is a contradiction, so there are no solutions to this system of equations. In set builder notation: \begin{equation*} \bigl\{ \bigr\} \end{equation*}. 
\newpage


## Problem 1B.

**Starting Equation**

$$
\begin{equation}
1x_1 - 7x_2 + 0x_3 + 6x_4 = 1 \\
1x_1 + 0x_2 - 2x_3 - 3x_4 = 0 \\
-1x_1 + 21x_2 - 4x_3 - 24x_4 = -3 \\
\end{equation}
$$

Convert this system of linear equations into an augmented matrix:

**Starting Matrix:** \

$$
\begin{bmatrix} 
1  & -7 & 0 & 6 & 1 \\ 
1 & 0 & -2 & -3 & 0 \\
-1 & 21 & -4 & -24 & -3 \\
\end{bmatrix}
$$

If I plug this matrix into Sage and run the `rref` function, like this:

```
B=Matrix(QQ,[[1,-7,0,6,1],[1,0,-2,-3,0],[-1,21,-4,-24,-3]])
print(B.rref())
```
I get: \
$$
\begin{bmatrix} 
1  & 0 & -2 & -3 & 0 \\ 
0 & 1 & -\frac{2}{7} & -\frac{9}{7} & -\frac{1}{7} \\
0 & 0 & 0 & 0 & 0 \\
\end{bmatrix}
$$
I don't see any clear contradictions, but we don't know what $x_3$ and $x_4$ are. If we convert to a system of linear equations again, we get:
$$
\begin{equation}
1x_1 + 0x_2 - 2x_3 - 3x_4 = 0 \\
0x_1 + 1x_2 - \frac{2}{7}x_3 - \frac{9}{7}x_4 = -\frac{1}{7} \\
0x_1 + 0x_2 - 0x_3 - 0x_4 = 0 \\
\end{equation}
$$
For clarity, we can drop the last equation, which doesn't actually give us any information (other than to tell us that 0 = 0), drop the terms with a 0 coefficient (since these also don't add any information). This leaves us with: \
$$
\begin{equation}
1x_1 - 2x_3 - 3x_4 = 0 \\
1x_2 - \frac{2}{7}x_3 - \frac{9}{7}x_4 = -\frac{1}{7} \\
\end{equation}
$$
We can move around some terms to get:
$$
\begin{equation}
1x_1 = 2x_3 + 3x_4 \\ 
1x_2   =  \frac{9}{7}x_4 + \frac{2}{7}x_3 -\frac{1}{7} \\
\end{equation}
$$
Therefore, the solution set is:

$$
\begin{equation} 
  \left\{ \left[
  \begin{matrix}
    2x_3 + 3x_4 \\ \frac{9}{7}x_4 + \frac{2}{7}x_3 -\frac{1}{7} \\ x_3\\ x_4
  \end{matrix}
  \right] \Bigg|  x_3, x_4 \in \mathbb{R}  \right\}
\end{equation} 
$$


\newpage
## Problem 1C.

**Starting Equation**


\begin{equation}
2x_1 - x_2 + 0x_3 + 0x_4 = 1 \\
-2x_1 + 0x_2 + x_3 + 0x_4 = 1 \\
x_1 + 0x_2 + 2 x_3 - 4x_4 = 1 \\
-2x_1 + 0x_2 + 2x_3 + -4x_4 = 1 \\
\end{equation}


Convert this to an augmented matrix:

**Starting Matrix:** \

$$
\begin{bmatrix} 
2 & -1 & 0 & 0 & 1 \\ 
-2 & 0 & 1 & 0 & 1 \\
1 & 0 & 2 & -4 & 1 \\
-2 & 0 & 2 & -4 & 1 \\
\end{bmatrix}
$$


If I plug this matrix into Sage and run the `rref` function, like this:

```
A=Matrix(QQ,[[2,-1,0,0,1],[-2,0,1,0,1],[1,0,2,-4,1],[-2,0,2,-4,1]])
print(A.rref())
```

$$
\begin{bmatrix} 
1 & 0 & 0 & 0 & 0 \\ 
0 & 1 & 0 & 0 & -1 \\
0 & 0 & 1 & 0 & 1 \\
0 & 0 & 0 & 1 & \frac{1}{4} \\
\end{bmatrix}
$$


If we put this back into a system of equations, we get:


\begin{equation}

x_1 = 0 \\
x_2 = -1 \\
x_3 = 1 \\
x_4 = \frac{1}{4} \\
\end{equation}



In set builder notation:


\begin{equation}

  \left\{ \left[
  \begin{matrix}
    0 \\ -1 \\ 1 \\ \frac{1}{4}
  \end{matrix}
  \right] \right\}
\end{equation}


\newpage


## Problem 2.

Starting equation:

\begin{equation}

x_1 - x_2 = 1 \\
3x_1 - 3x_2 = k \\
\end{equation}



Convert to an augmented matrix:

**Starting Matrix:** \

$$
\begin{bmatrix} 
1 & -1 & 1  \\ 
3 & -3 & k  \\
\end{bmatrix}
$$

We might not be able to get this exactly into RREF, but we can try to get as close as we can:

```{r include = F}
i = 0
```
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 2 with the difference between row 2 and 3 times row 1 $ **(R2 = R2 - 3R1)** \
$$
\begin{bmatrix} 
1 & -1 & 1  \\ 
0 & 0 & k-3  \\
\end{bmatrix}
$$

## Problem 2A. 
If we want this matrix to be in RREF, one potential option is for `k-3` to be equal to 1, in which case `k = 4`. Then, our matrix would look like this:

$$
\begin{bmatrix} 
1 & -1 & 1  \\ 
0 & 0 & 1  \\
\end{bmatrix}
$$

But we aren't in RREF just yet, we have a 1 above our new pivot on row 2. Let's fix that by subtracting row 2 from row 1 (R1 = R1 - R2).
$$
\begin{bmatrix} 
1 & -1 & 0  \\ 
0 & 0 & 1 \\
\end{bmatrix}
$$
Let's turn it back into a system of equations:


\begin{equation}

x_1 - x_2 = 0\\
0 - 0 = 1\\
\end{equation}



**We have a contradiction in the second equation, so this describes the scenario in problem 2a where the system is inconsistent.**
In set notation, 


\begin{equation}

  \left\{ \right\}
\end{equation}


## Problem 2C
If we want this matrix to be in RREF, one potential option is for `k-3` to be equal to 0, in which case `k = 3`. Then our matrix is in RREF already (bottom row would be all zeroes) and looks like this when we turn it into a system of equations:


\begin{equation}

x_1 - x_2 = 1 \\
0 - 0 = 0 \\
\end{equation}


We could re-phrase the top equation as $x_1 = x_2 + 1$, to describe $x_1$ in terms of $x_2$ 

And describe this using set notation setting $x_2$ to `a`:


\begin{equation}

  \left\{ \left[
  \begin{matrix}
    a+1 \\ a 
  \end{matrix}
  \right] \Bigg|  a \in \mathbb{R}  \right\}
\end{equation}



*This describes the scenario in problem 2c, where we have infinitely many solutions and `k = 3`*.

## Problem 2B.
This leaves only the scenario described in problem 2b where we want to find a solution set consistent with a unique solution. My intuition here is that this isn't possible because we don't have enough information to find a unique solution. Perhaps if we have one more hypothetical equation it would help. Let's set k = 3 and add another equation to the mix.


\begin{equation}

x_1 - x_2 = 1 \\
3x_1 - 3x_2 = 3 \\
x_1 + x_2 = 7 \\
\end{equation}



As a matrix:

$$
\begin{bmatrix} 
1 & -1 & 1  \\ 
3 & -3 & 3 \\
1 & 1 & 7 \\
\end{bmatrix}
$$
The matrix in RREF form:
$$
\begin{bmatrix} 
1 & 0 & 4  \\ 
0 & 1 & 3 \\
0 & 0 & 0 \\
\end{bmatrix}
$$

Then convert it back to a system of equations:


\begin{equation}

x_1 = 4 \\
x_2 = 3 \\
0 + 0 = 0 \\
\end{equation}

What is the difference between these three outcomes? Recall that **Definition 1.4.2** in `https://teambasedinquirylearning.github.io/linear-algebra/2022/pdf/main.pdf` states that the pivot columns in an augmented matrix correspond to the bound variables of a system of equations, while the remaining variables are called free variables.

To summarize:

When k = 4, there are zero solutions and the RREF has one free variable and two pivots/bound variables (indicated with asterisks), one of which is in the last column, which makes a contradiction and an inconsistent system.
$$
\begin{bmatrix} 
*1* & -1 & 0  \\ 
0 & 0 & *1*  \\
\end{bmatrix}
$$
When k = 3, there are infinitely many solutions and the RREF has one pivot/bound variable (indicated with asterisks) and one free variable (excluding the column on the right side of the matrix, which might actually be considered a constant in this context - I'm not totally sure):
$$
\begin{bmatrix} 
*1* & -1 & 1  \\ 
0 & 0 & 0  \\
\end{bmatrix}
$$

When k = 3 and we add another equation to the mix, we get a unique solution and the RREF has two bound variables/pivots (indicated with asterisks) and no free variable (again, excluding whats in the last column).

$$
\begin{bmatrix} 
*1* & 0 & 4  \\ 
0 & *1* & 3 \\
0 & 0 & 0 \\
\end{bmatrix}
$$

I hypothesize that:
- when we have a consistent system with `n` unknowns and `n-1` bound variables with one free variable then we have infinite solutions.
- when we have a consistent system with `n` unknowns and `n` bound variables with no free variable then we have a single solution.

\newpage

## Problem 3
$$
\begin{bmatrix} 
1 & -3 & 2 & g \\ 
2 & -1 & 3 & h  \\
3 & -4 & 5 & k  \\
\end{bmatrix}
$$

What values could g, h, and k assume while still having a consistent system? Let's try to get this into RREF as much as possible:


```{r include = F}
i = 0
```
```{r include = F}
i = i + 1
```
**Step `r i`**. Replace row 2 with the difference between row 2 and 2 times row 1 (R2 = R2 - 2R1)
```{r include = F}
i = i + 1
```

$$
\begin{bmatrix} 
1 & -3 & 2 & g \\ 
0 & 5 & -1 & h-2g  \\
3 & -4 & 5 & k  \\
\end{bmatrix}
$$


**Step `r i`**. Replace row 3 with row 3 - 3 times row 1 (R3 = R3 - 3R1)
```{r include = F}
i = i + 1
```
$$
\begin{bmatrix} 
1 & -3 & 2 & g \\ 
0 & 5 & -1 & h-2g  \\
0 & 5 & -1 & k-3g  \\
\end{bmatrix}
$$


**Step `r i`**. Replace row 3 with row 3 - row 2 (R3 = R3 - R2).
```{r include = F}
i = i + 1
```
$$
\begin{bmatrix} 
1 & -3 & 2 & g \\ 
0 & 5 & -1 & h-2g  \\
0 & 0 & 0 & k-h-g \\
\end{bmatrix}
$$

**Step `r i`**. Replace row 2 with row 2 times 1/5 (R2 = 1/5 R2).
```{r include = F}
i = i + 1
```
$$
\begin{bmatrix} 
1 & -3 & 2 & g \\ 
0 & 1 & -\frac{1}{5} & \frac{1}{5}h - \frac{2}{5}g  \\
0 & 0 & 0 & k-h-g \\
\end{bmatrix}
$$

**Step `r i`**. Replace row 1 with row 1 + 3 times row 2 (R1 = R1 + 3R2) to get 0s in pivot column.
```{r include = F}
i = i + 1
```
$$
\begin{bmatrix} 
1 & 0 & \frac{7}{5} & \frac{3}{5}h - \frac{1}{5}g \\ 
0 & 1 & -\frac{1}{5} & \frac{1}{5}h - \frac{2}{5}g  \\
0 & 0 & 0 & k-h-g \\
\end{bmatrix}
$$
There are pivots in the columns for $x_1$ and $x_2$, but $x_3$ is free, so there are infinite solutions to this system. 

Without more concrete information about the other variables ($x_1$, $x_2$, and $x_3$), I don't think there's much more we can say other than:



\begin{equation}

  \left\{ \left[
  \begin{matrix}
    g \\ h \\ k \\x_1 \\x_2 \\x_3
  \end{matrix}
  \right] \Bigg|  g,h,k,x_,x_2,x_3 \in \mathbb{R}, k-h-g=0,     x_1 + \frac{7}{5}x_3 = \frac{3}{5}h - \frac{1}{5}g , x_2 - \frac{1}{5}x_3 = \frac{1}{5}h - \frac{2}{5}g \right\}
\end{equation}
  




For example, if $x_1 = 2$ and $x_2 = 4$, but $x_3 = 6$, then 

$$
\begin{bmatrix} 
2 & -12 & 12 & g \\ 
4 & -4 & 18 & h  \\
6 & -16 & 30 & k  \\
\end{bmatrix}
$$
`g`= 2, `h` = 18, `k` = 20.

$k-h-g = 0$ becomes $20 - 18 - 2 = 0$.

$x_1 + \frac{7}{5}x_3 = \frac{3}{5}h - \frac{1}{5}g$ becomes $2 + \frac{7}{5}\times6 = \frac{3}{5}\times18 - \frac{1}{5}\times2$ when we plug in, which simplifies to $\frac{52}{5} = \frac{52}{5}$

$x_2 - \frac{1}{5}x_3 = \frac{1}{5}h - \frac{2}{5}g$ becomes $ \frac{20}{5} - \frac{6}{5} = \frac{18}{5} - \frac{4}{5}$, which simplifies to  $\frac{14}{5} = \frac{14}{5}$


\newpage

# Problem 4
Construct a system of three linear equations in three variables having:

(a) 0 free variables;

(b) 1 free variable;

(c) 2 free variables. 

In each case, solve the system you created and write the solution set using set-builder notation. Make a conjecture about the relationship between the number of free variables in a linear system and the type of solution set that can be obtained from a given system. 

0 free variables:
$$
\begin{bmatrix} 
1 & 0 & 0 & 1 \\ 
0 & 1 & 0 & 2 \\
0 & 0 & 1 & 3 \\
\end{bmatrix}
$$
Solution set:

\begin{equation}

  \left\{ \left[
  \begin{matrix}
    1 \\ 2 \\ 3
  \end{matrix}
  \right] \right\}
\end{equation}




(b) 1 free variable;

$$
\begin{bmatrix} 
1 & 0 & 0 & 1 \\ 
0 & 1 & 0 & 2 \\
0 & 0 & 0 & 0 \\
\end{bmatrix}
$$

Solution set:

\begin{equation}

  \left\{ \left[
  \begin{matrix}
    1 \\ 2 \\ a  
  \end{matrix}
  \right] \Bigg|  a \in \mathbb{R}  \right\}
\end{equation}



(b) 2 free variables;

$$
\begin{bmatrix} 
1 & 0 & 0 & 1 \\ 
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
\end{bmatrix}
$$

Solution set:

\begin{equation}

  \left\{ \left[
  \begin{matrix}
    1 \\ a \\ b  
  \end{matrix}
  \right] \Bigg|  a,b \in \mathbb{R}  \right\}
\end{equation}


I'm going to revise my hypothesis from problem 2 a little bit: 
- when we have a consistent system with `n` unknowns and `n` bound variables with no free variable then we have a single solution.
- when we have a consistent system with `n` unknowns and one or more free variables then we have infinite solutions.
- if we have a pivot in the last column of the augmented matrix, we have a contradiction and the system is inconsistent (because it means 0=1). 

\newpage


## Problem 5A
**If a system of equations having four equations and three unknowns is consistent, then the solution is unique.**
Not necessarily. Similarly to what was happening in problem 2, additional equations can be equivalent in terms of the relationships between the different variables that are being represented, in which case you could have really any number of equations - the free and bound variables are what is important.

## Problem 5B
**Suppose that a linear system has five equations and three unknowns and that the RREF of the coefficient matrix has a pivot in every column. Then the linear system is consistent and has a unique solution.**
If you have 3 unknowns and a pivot in all 3 columns corresponding to the unknowns coefficients, then you would have 3 bound variables and a unique solution. However, if this is an augmented matrix and 'pivot in every column' includes the last column, then the system is inconsistent and has zero solutions.

## Problem 5C
**Suppose that a linear system has four equations and four unknowns and that the RREF of the coefficient matrix has four pivots. Then the linear system is consistent and has a unique solution.**
If you have 4 unknowns and 4 pivots, then you would have 4 bound variables and a unique solution. However, if this is an augmented matrix and 'pivot in every column' includes the last column, then the system is inconsistent and has zero solutions.



